---
title: "Final Project"
author: "Zhuoyu"
date: "2022-12-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load packages
```{r}
rm(list=ls())
library(tidyverse)
library(DBI)
library(dbplyr)
library(dplyr)
library(bigrquery)
library(gapminder) # import data gapminder
library(ggplot2)
```
## import data
### connect
```{r}
#bq_deauth()
project <- "secret-griffin-368601"
con <- dbConnect(
  bigrquery::bigquery(), #signifying dbConnect that you want to connect to a BigQuery data source
  project = "bigquery-public-data",
  dataset = "world_bank_health_population",
  billing = project
)
con 
dbListTables(con)
```


### select variables interested

The names of variables can be found here : https://datatopics.worldbank.org/health/available-indicators

- medical resources：
-     Hospital beds (per 1,000 people) 
-     Nurses and midwives (per 1,000 people)
-     Physicians (per 1,000 people)
- the prevalence of diseases
-     Prevalence of HIV, total (% of population ages 15-49)
-     Incidence of malaria (per 1,000 population at risk)
- health expenditure
-     Current health expenditure per capita (current US$)
- Population:Population growth (annual %)

#### create sql object
```{r}
sql<-"
SELECT country_name, indicator_name,value, year
FROM health_nutrition_population
WHERE 
      indicator_name =  'Hospital beds (per 1,000 people)' OR
      indicator_name =  'Nurses and midwives (per 1,000 people)' OR
      indicator_name =  'Physicians (per 1,000 people)' OR
      indicator_name =  'Prevalence of hypertension (% of adults ages 30-79)' OR
      indicator_name =  'Incidence of malaria (per 1,000 population at risk)' OR
      indicator_name =  'Current health expenditure per capita (current US$)' OR
      indicator_name = 'Population growth (annual %)' OR
      indicator_name = 'Prevalence of HIV, total (% of population ages 15-49)'
        "
```

```{r}
#bigquery data saved as a R object
Robject<-dbGetQuery(con, sql)

#long format -> wide format
data1<- spread(Robject, 
               key = indicator_name, 
               value = value) 

#import variables regions
data <- readRDS("healthexp.Rds")

# change first column name into "Country"
colnames(data1)[1] = "Country"

#combine Region with country
dta <- data1 %>% left_join(data[,c(1,2)], by = "Country")

#droping rows with missing values
dta_complete<-na.omit(dta)
dta_complete<-dta_complete%>%
  filter(year != 2018)

#removing duplicate rows
dta_complete_unique <- dta_complete[!duplicated(dta_complete), ]
#remove other datasets
colnames(dta_complete_unique)[3:9] <-c("Health expenditure per capita(US)",
  "Hospital beds(per 1000 people)",
  "Incidence of Malaria(per 1000 population)",
  "Number of nurses(per 1000 people)",
  "Number of physicians(per 1000 people)",
  "Population growth(annual percent)",
  "Prevalence of HIV(percent)")

write_csv(dta_complete_unique, file = "dta_processed_update2.csv")


```


## disconnect 
```{r}
dbDisconnect(con)
```

